Topological Equivariant Artist Model for Visualization Library Architecture

### 2022-07-12 paper/main.tex
working references

\IEEEPARstart{V}isualizations are expected to be a faithful translation of the inherent structure of the data being visualized; therefore the components of visualization software libraries that translate data to graphics are also expected to preserve this structure. Motivated by a need to develop new library components that adapt to modern data needs, especially high dimensional multivariate and distributed data, we propose a framework that topological and field structure of the input data that must be mapped to the visualization and which changes to data must be expressed in visualization, and a a functional design framework that describes composability in terms of functional algebraic operations. \note{this may be a good place to incorporate types} This allows complex visualizations to be built from simple verifiable parts\cite{huHowFunctionalProgramming2015,hughesWhyFunctionalProgramming1989}, which makes these components easier to drop into the existing library and be used by domain specific library developers to build new libraries that work with existing ones.  While many of these elements have been partially described elsewhere, our framework naturally combines them in a unified framework based on ideas in category theory. Further we see the application of these ideas in low level visualization library design as our motivation, and we hope these ideas will help shape the reorganization of critical data visualization libraries, such as Matplotlib. The contribution of this paper is a uniform method for describing the structure in data and a mathematical framework for describing, constructing, and testing structure preserving visualization library components.

@@

There is an assumption in visualization that the connectivity of the elements in a dataset are preserved in the visual representation; roughly topology provides a method for encoding this connectivity \cite{wilkinsonGrammarGraphics2005}. Bertin gives the examples of discrete data values should be represented by discrete entities such as points, quantitive continuous data by lines, and surfaces through area marks \cite{bertinSemiologyGraphicsDiagrams2011a}. In \autoref{fig:related-work:continuity:ktypes} the records of the GHCN\cite{lawrimoreGlobalHistoricalClimatology2011} weather station table can be considered disconnected from each other and therefore have a 0D continuity. This means each distinct row can be reduced to a discrete point and it is expected  that each rows will be encoded as a discrete visual element, such as a standalone scatter point. Each station can also be considered a time series of weather observation since temperature and pressure are sampled from a 1D continuous measurement space. The connectivity of these observations can be modeled as an interval and the data is encoded as a line since that visual element is also 1D continuous. The measurements at set of stations on any given day is a sparse sampling of a 2D geospatial grid and the connectivity can be modeled as a 2D plane. Instead of interpolation, which is how continuity was preserved in the timeseries plot, in this instance the connectivity is preserved by plotting the stations in the map such that their relative distances are preserved. Encoding connectivity using the formalism of topological spaces allows us to encode the connectivity of the data in a uniform manner such that we can then generalize the preservation of that connectivity in the visual representation rather than constructing a condition specifically for each type of topology.

@@

The topological structure of the data can be expressed programmatically by constructing data types that encapsulate the class of topological structure-.i.e point, line, plane, network, etc. We propose that these data types can be formally specified as objects of a category. 
\begin{definition} 
   An arbitrary \textbf{category} $\mathcal{C}$ consists of the following \textit{data}: 
\begin{enumerate}
  \item \textit{object}: X
  \item \textit{identity morphism}: $X \xrightarrow {id_x} X$
  \item \textit{morphisms}: $X \xrightarrow{f} Y$ between objects X, Y in $\mathcal{C}$
  \item \textit{composite morphism}: given the morphisms $X \xrightarrow{f} Y$ and $Y \xrightarrow{g} Z$, there exists a morphism $X \xrightarrow{g\circ f} Z$
\end{enumerate}
The data must satisfy the following \textit{rules}
\begin{enumerate}
  \item \textit{unitality:} for every morphism $ X \xrightarrow{f} Y$, there exists identity morphisms $id_x, id_y$ such that $f \circ id_x = f = id_y \circ f$
  \item \textit{associativity:} if any three morphisms $f, g, h$ are composable, 
    \begin{equation*}
      \begin{tikzcd}
        X \arrow[r, "f"] \arrow[rrr, "h\circ(g\circ f) = (h\circ g)\circ f"', bend right, dashed] & Y  \arrow[r, "g"] & Z \arrow[r, "h"] & W
        \end{tikzcd}}
  \end{equation*}
  then they are associative such that $h\circ(g\circ f) = (h \circ g) \circ f$  \cite{lawvere2009conceptual,riehlCategoryTheoryContext,maclaneCategoriesWorkingMathematician2013,fongInvitationAppliedCategory2019}. 
  \end{enumerate}
\end{definition}

@@

\begin{definition} A \textbf{type specification} is the bundle map $\pi: \mathscr{U} \rightarrow \textbf{DT}$. The base space $\textbf{DT}$ is the set of data types, and the total space $\mathscr{U}$ is the set of domains of those data types. Given an arbitrary data type $T \in \textbf{DT}$, the preimage $\pi^{-1} \subset \mathscr{U}$ is the domain of $T$ and $x \in \pi^{-1}(T)$ is an object of type $T$.  
\end{definition}
For example, if $T=\texttt{int}$, then the image $\pi^{-1}(\texttt{int}) = \mathbb{Z} \subset \mathscr{U}$ is the set of all integers. The domain $\mathscr{U}_{{\sigma}}$ is also a topological space, meaning there is no restriction on the dimensionality of the field, thus allowing for the encoding of multi-dimensional fields in the same manner as single dimensional fields. Since many fields can have the same datatype, Spivak formally defines a mapping from field name to field data type, akin to a  data base schema \cite{ullmanFirstCourseDatabase2008}. 
\begin{definition} A \textbf{schema} consists of a pair $(\fnames, \sigma)$ where $\fnames$ is the set of field names and $\sigma: \fnames \rightarrow \textbf{DT}$ is a function from field name to field data type. 
\end{definition}
The function $\sigma$ is composed with $\pi$ such that the bundle  $\pi_{\sigma}:\mathscr{U}_{\sigma} \rightarrow \fnames$ associates a field name $c \in C$ with the corresponding data type domain. Spivak then defines data records as sections on the bundle defined by $\pi_{\sigma}$. 
\begin{definition} A \textbf{record} is a function $\delement: \fnames \rightarrow \mathscr{U}_{\sigma}$ and the set of records on $\pi_{\sigma}$ is denoted $\Gamma^{\pi}(\sigma)$. Records must return an object of type $\sigma(\fname) \in \textbf{DT}$ for each field $c \in C$.
\end{definition}
Spivak then constructs tables as sections $\dsection: \dbase \rightarrow \Gamma^{\pi}(\sigma)$ from a key space $\dspace$ to the set of records $\Gamma^{\pi}(\sigma)$ on the schema bundle. Generalizing Spivak's formulation of tables, we define the fiber space in this work as the set of record functions 
\begin{equation}
  \dfiberc \coloneqq \{\delement: \fnames \rightarrow \mathscr{U}_{\sigma} \bigm{\vert} \pi_{\sigma}(\delement(\fname)) = \fname\;for\;all\; \fname \in \fnames \}
\end{equation}
such that the preimage of a point is the corresponding data type domain $\pi^{-1}(\dbasepoint) = \dfiber_{k} = \mathscr{U}_{{\sigma}_{\dbasepoint}}$. The construction of the fiber as the space of records provides a way to distinguish data fields with the same datatype while allowing that datatype to be defined as almost any arbitrary topological space.

@@

The presheaf preserves continuity by stating that given a morphism $\iota$ between two objects of $\mathcal{\dbase}$ 
\begin{equation*}
  \iota: \openset_1 \rightarrow \openset_2
\end{equation*}
there must be a corresponding restriction morphiism $\iota^*$ between the sets of sections 
\begin{equation*}
  \iota^*: \cgamma{\openset_2}{\dtotal_{\openset_2}} \rightarrow \cgamma{\openset_1}{\dtotal_{\openset_1}}
\end{equation*}
which are the outputs of the sheaf $\sheaf(\openset) = \cgamma{\openset}{\dtotal\restriction_{\openset}}$. The presheaf is contravariant because the morphisms between the input openset objects go in the opposite direction from the morphisms between the output set objects. The expectation that data containers implement inclusion and restriction morphisms means that, for example, if a data container implements subsetting (inclusion) of a continuous base space than it must ensure that the sections can be interpolated such that they can be defined over the subspace.

@@

\subsection{Equivariance}
\label{sec:artist:equiv}
Equivariance in a visual context was formally defined as the preservation of a binary operator from one domain to another by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987}. We generalize binary operations to actions $\dfunc \in \dfuncset$ on the sheaf $\sheaf_{\dbase, \dtotal}$. 
\begin{definition} An \textbf{action}  $\act: \dfuncsetc \times \dtotalc \rightarrow \sheaf_{\dbase, \dtotal}$ such that 
  \begin{equation}
    act(\dfuncc_{j}, act(\dfuncc_{i}, \dsectionc)) = act(\dfuncc_{j} \circ \dfuncc_{i}, \dsectionc)
  \end{equation}
where $\dfunc_{i}, \dfunc_{j}$ are elements of the set $\dfuncset$ equipped with operator $\circ$ and \dsection\ is an element of the sheaf  $\sheaf_{\dbase, \dtotal}$.\cite{nlab:action}
\end{definition}.

@@

As introduced in \autoref{def:atct:io:fiber}, the fiber transformation $\dfunct: \dfuncpull \dtotal_{\dbasepoint^{\prime}} \rightarrow \dfuncpull \dtotal_{\dbasepoint^{\prime}}$ is a morphism on the fiber $\dfunct \in Hom(\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}},\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}})$ restricted to a point $\dbasepoint^{\prime} \in \openset^{\prime}$. The fiber transformation is a change in section 
\begin{equation}
  \label{eq:atct:morphism:fiber}
  \dfunct: \dfuncpull \dsection \restriction_{\openset^{\prime}} \mapsto \dfuncpull \dsection^{\prime} \restriction_{\openset}
\end{equation}
where $\dsection, \dsection^{\prime} \in \Gamma(\openset^{\prime}, \dfuncpull\dtotal\restriction_{\openset^{\prime}})$. Since there are many different fiber transformations,  there is not a uniform way of expressing how this morphism acts on sections unlike in \label{eq:atct:morphism:verify_base}. Examples of a fiber transformation are actions such as partial ordering for ranking purpose \cite{bruggemannRankingPrioritizationMultiindicator2011} and the permutation, ordering, translation, and scaling codified as Steven's measurement scales \cite{stevensTheoryScalesMeasurement1946m, leaFormalizationMeasurementScale, thomasMathematizationNotMeasurement2014}.

@@

The difference between $\dtotal$ and $\vtotal$ are semantic rather than structural; they are both maps from the topological space \dbase\ to sets of functions that return records of values. This means any $\vtotal$ can be redefined as $\dtotal$, 
\begin{equation}
  \label{eq:construction:nu:fabrication}
  \begin{tikzcd}[col sep=Huge]
    \dfiberc_{\dbasepointc} 
    \arrow[rr, "\vchannelc", color=artist] 
    \arrow[rrrr, "\vchannelc^{\prime\prime}", dashed, bend right, color=artist] &  & 
    \vfiberc_{\dbasepointc}\coloneqq{\dfiberc_{\dbasepointc}^{\prime}} 
    \arrow[rr, "\vchannelc^{\prime}", color=artist] &  & 
    \vfiberc^{\prime}_{\dbasepointc}
  \end{tikzcd}
\end{equation}
which means that, as shown in \autoref{eq:construction:nu:fabrication}, any collection of $\vchannel$ functions can be composed such that they are equivalent to a $\vchannel$ that directly converts the input to the output. As with artists, $\vchannel$ are maps of sections such that the composition operators defined in \autoref{sec:artist:composition} can also act on transformers $\vchannel$, meaning that encoders can be added $\vchannel_{a+b} = \vchannel_{a} + \vchannel_{b}$ and multiplied d $\vchannel_{a\times b} = \vchannel_{a}  \vchannel_{b}$.  Encoders designed to satisfy these composability constraints provide for a rich set of building blocks for implementing complex encoders.

@@

    def draw(self, renderer, bounds, rate):
        for tau_local in self.section.query(bounds, rate): 
            mu = self.artist.nu(tau_local)
            rho = self.artist.qhat(**mu)
            output = rho(renderer)
\end{minted}
As shown in the \texttt{draw} method, generating a graphic section $\gsection$ is implemented as the composition of $\texttt{qhat} \approx \vmark$ and $\texttt{nu} \approx \vchannel$ applied to a local section of the sheaf $\textttt{self.section.query} \approx \desection^{i}$ such  $\texttt{draw} \approx \vmarkc\circ\vchannelc\circ\dsectionc  = \vartistc\circ \dsectionc$. The $\vchannel$ and $\vmark$ functions shown here are written such that they can generate a visual element given a local section $\dsection\restriction_{\dbase^{i}}$ which can be as little or large as needed. This flexibility is a prerequisite for building scalable and streaming visualizations that may not have access to all the data.

@@

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
\bibliography{bibliography}

>>>>

\IEEEPARstart{V}isualizations are expected to be a faithful translation of the inherent structure of the data being visualized; therefore the components of visualization software libraries that translate data to graphics are also expected to preserve this structure. Motivated by a need to develop new library components that adapt to modern data needs, especially high dimensional multivariate and distributed data, we propose a framework that topological and field structure of the input data that must be mapped to the visualization and which changes to data must be expressed in visualization, and a a functional design framework that describes composability in terms of functional algebraic operations. \note{this may be a good place to incorporate types} This allows complex visualizations to be built from simple verifiable parts\cite{huHowFunctionalProgramming2015, hughesWhyFunctionalProgramming1989}, which makes these components easier to drop into the existing library and be used by domain specific library developers to build new libraries that work with existing ones.  While many of these elements have been partially described elsewhere, our framework naturally combines them in a unified framework based on ideas in category theory. Further we see the application of these ideas in low level visualization library design as our motivation, and we hope these ideas will help shape the reorganization of critical data visualization libraries, such as Matplotlib. The contribution of this paper is a uniform method for describing the structure in data and a mathematical framework for describing, constructing, and testing structure preserving visualization library components.

@@

There is an assumption in visualization that the connectivity of the elements in a dataset are preserved in the visual representation; roughly topology provides a method for encoding this connectivity \cite{wilkinsonGrammarGraphics2005}. Bertin gives the examples of discrete data values should be represented by discrete entities such as points, quantitative continuous data by lines, and surfaces through area marks \cite{bertinSemiologyGraphicsDiagrams2011}. In \autoref{fig:related-work:continuity:ktypes} the records of the GHCN\cite{lawrimoreGlobalHistoricalClimatology2011} weather station table can be considered disconnected from each other and therefore have a 0D continuity. This means each distinct row can be reduced to a discrete point and it is expected  that each rows will be encoded as a discrete visual element, such as a standalone scatter point. Each station can also be considered a time series of weather observation since temperature and pressure are sampled from a 1D continuous measurement space. The connectivity of these observations can be modeled as an interval and the data is encoded as a line since that visual element is also 1D continuous. The measurements at set of stations on any given day is a sparse sampling of a 2D geospatial grid and the connectivity can be modeled as a 2D plane. Instead of interpolation, which is how continuity was preserved in the timeseries plot, in this instance the connectivity is preserved by plotting the stations in the map such that their relative distances are preserved. Encoding connectivity using the formalism of topological spaces allows us to encode the connectivity of the data in a uniform manner such that we can then generalize the preservation of that connectivity in the visual representation rather than constructing a condition specifically for each type of topology.

@@

The topological structure of the data can be expressed programmatically by constructing data types that encapsulate the class of topological structure-.i.e point, line, plane, network, etc. We propose that these data types can be formally specified as objects of a category. 
\begin{definition} 
   An arbitrary \textbf{category} $\mathcal{C}$ consists of the following \textit{data}: 
\begin{enumerate}
  \item \textit{object}: X
  \item \textit{identity morphism}: $X \xrightarrow {id_x} X$
  \item \textit{morphisms}: $X \xrightarrow{f} Y$ between objects X, Y in $\mathcal{C}$
  \item \textit{composite morphism}: given the morphisms $X \xrightarrow{f} Y$ and $Y \xrightarrow{g} Z$, there exists a morphism $X \xrightarrow{g\circ f} Z$
\end{enumerate}
The data must satisfy the following \textit{rules}
\begin{enumerate}
  \item \textit{unitality:} for every morphism $ X \xrightarrow{f} Y$, there exists identity morphisms $id_x, id_y$ such that $f \circ id_x = f = id_y \circ f$
  \item \textit{associativity:} if any three morphisms $f, g, h$ are composable, 
    \begin{equation*}
      \begin{tikzcd}
        X \arrow[r, "f"] \arrow[rrr, "h\circ(g\circ f) = (h\circ g)\circ f"', bend right, dashed] & Y  \arrow[r, "g"] & Z \arrow[r, "h"] & W
        \end{tikzcd}
  \end{equation*}
  then they are associative such that $h\circ(g\circ f) = (h \circ g) \circ f$  \cite{lawvere2009conceptual,riehlCategoryTheoryContext,maclaneCategoriesWorkingMathematician2013,fongInvitationAppliedCategory2019}. 
  \end{enumerate}
\end{definition}

@@

\begin{definition} A \textbf{type specification} is the bundle map $\pi: \mathscr{U} \rightarrow \textbf{DT}$. The base space $\textbf{DT}$ is the set of data types, and the total space $\mathscr{U}$ is the set of domains of those data types. Given an arbitrary data type $T \in \textbf{DT}$, the preimage $\pi^{-1} \subset \mathscr{U}$ is the domain of $T$ and $x \in \pi^{-1}(T)$ is an object of type $T$.  
\end{definition}
For example, if $T=\texttt{int}$, then the image $\pi^{-1}(\texttt{int}) = \mathbb{Z} \subset \mathscr{U}$ is the set of all integers. The domain $\mathscr{U}_{{\sigma}}$ is also a topological space, meaning there is no restriction on the dimensionality of the field, thus allowing for the encoding of multi-dimensional fields in the same manner as single dimensional fields. Since many fields can have the same datatype, Spivak formally defines a mapping from field name to field data type, akin to a  data base schema \cite{ullmanFirstCourseDatabase2008}. 
\begin{definition} A \textbf{schema} consists of a pair $(\fnames, \sigma)$ where $\fnames$ is the set of field names and $\sigma: \fnames \rightarrow \textbf{DT}$ is a function from field name to field data type. 
\end{definition}
The function $\sigma$ is composed with $\pi$ such that the bundle  $\pi_{\sigma}:\mathscr{U}_{\sigma} \rightarrow \fnames$ associates a field name $c \in C$ with the corresponding data type domain. Spivak then defines data records as sections on the bundle defined by $\pi_{\sigma}$. 
\begin{definition} A \textbf{record} is a function $\delement: \fnames \rightarrow \mathscr{U}_{\sigma}$ and the set of records on $\pi_{\sigma}$ is denoted $\Gamma^{\pi}(\sigma)$. Records must return an object of type $\sigma(\fname) \in \textbf{DT}$ for each field $c \in C$.
\end{definition}
Spivak then constructs tables as sections $\dsection: \dbase \rightarrow \Gamma^{\pi}(\sigma)$ from a key space $\dbase$ to the set of records $\Gamma^{\pi}(\sigma)$ on the schema bundle. Generalizing Spivak's formulation of tables, we define the fiber space in this work as the set of record functions 
\begin{equation}
  \dfiberc \coloneqq \{\delement: \fnames \rightarrow \mathscr{U}_{\sigma} \bigm{\vert} \pi_{\sigma}(\delement(\fname)) = \fname\;for\;all\; \fname \in \fnames \}
\end{equation}
such that the preimage of a point is the corresponding data type domain $\pi^{-1}(\dbasepoint) = \dfiber_{k} = \mathscr{U}_{{\sigma}_{\dbasepoint}}$. The construction of the fiber as the space of records provides a way to distinguish data fields with the same datatype while allowing that datatype to be defined as almost any arbitrary topological space.

@@

The presheaf preserves continuity by stating that given a morphism $\iota$ between two objects of $\mathcal{\dbase}$ 
\begin{equation*}
  \iota: \openset_1 \rightarrow \openset_2
\end{equation*}
there must be a corresponding restriction morphiism $\iota^*$ between the sets of sections 
\begin{equation*}
  \iota^*: \cgamma{\openset_2}{\dtotal_{\openset_2}} \rightarrow \cgamma{\openset_1}{\dtotal_{\openset_1}}
\end{equation*}
which are the outputs of the sheaf $\sheaf(\openset) = \cgamma{\openset}{\dtotal\restriction_{\openset}}$. The presheaf is contravariant because the morphisms between the input openset objects go in the opposite direction from the morphisms between the output set objects. The expectation that data containers implement inclusion and restriction morphisms means that, for example, a data container mist ensure that sections can be interpolated over any subset that the container can provide.

@@

\subsection{Equivariance}
\label{sec:artist:equiv}
Equivariance in a visual context was formally defined as the preservation of a binary operator from one domain to another by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987}. We generalize binary operations to actions $\dfunc \in \dfuncset$ on the sheaf $\sheaf_{\dbase, \dtotal}$.

\begin{definition} 
  An \textbf{action}  $act: \dfuncsetc \times \dtotalc \rightarrow \sheaf_{\dbase, \dtotal}$ is a function that expresses that elements of set $\dfuncset$ act on elements of the sheaf of sections.Actions are associative such that 
  \begin{equation}
    act(\dfuncc_{j}, act(\dfuncc_{i}, \dsectionc)) = act(\dfuncc_{j} \circ \dfuncc_{i}, \dsectionc)
  \end{equation}
 for all elements  $\dfunc_{i}, \dfunc_{j}$ of the set $\dfuncset$ equipped with operator $\circ$ and all elements \dsection\ of the sheaf  $\sheaf_{\dbase, \dtotal}$. \cite{nlab:action}
\end{definition}.

@@

As introduced in \autoref{def:atct:io:fiber}, the fiber transformation $\dfunct: \dfuncpull \dtotal_{\dbasepoint^{\prime}} \rightarrow \dfuncpull \dtotal_{\dbasepoint^{\prime}}$ is a morphism on the fiber $\dfunct \in Hom(\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}},\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}})$ restricted to a point $\dbasepoint^{\prime} \in \openset^{\prime}$. The fiber transformation is a change in section 
\begin{equation}
  \label{eq:atct:morphism:fiber}
  \dfunct: \dfuncpull \dsection \restriction_{\openset^{\prime}} \mapsto \dfuncpull \dsection^{\prime} \restriction_{\openset}
\end{equation}
where $\dsection, \dsection^{\prime} \in \Gamma(\openset^{\prime}, \dfuncpull\dtotal\restriction_{\openset^{\prime}})$. Since there are many different fiber transformations,  there is not a uniform way of expressing how this morphism acts on sections unlike in \autoref{eq:atct:morphism:verify_base}. Examples of a fiber transformation are actions such as partial ordering for ranking purpose \cite{bruggemannRankingPrioritizationMultiindicator2011} and the permutation, ordering, translation, and scaling codified as Steven's measurement scales \cite{stevensTheoryScalesMeasurement1946m, leaFormalizationMeasurementScale, thomasMathematizationNotMeasurement2014}.

@@

The difference between $\dtotal$ and $\vtotal$ are semantic rather than structural; they are both maps from the topological space \dbase\ to sets of functions that return records of values. This means any $\vtotal$ can be redefined as $\dtotal$, 
\begin{equation}
  \label{eq:construction:nu:fabrication}
  \begin{tikzcd}[col sep=huge]
    \dfiberc_{\dbasepointc} 
    \arrow[rr, "\vchannelc", color=artist] 
    \arrow[rrrr, "\vchannelc^{\prime\prime}", dashed, bend right, color=artist] &  & 
    \vfiberc_{\dbasepointc}\coloneqq{\dfiberc_{\dbasepointc}^{\prime}} 
    \arrow[rr, "\vchannelc^{\prime}", color=artist] &  & 
    \vfiberc^{\prime}_{\dbasepointc}
  \end{tikzcd}
\end{equation}
which means that, as shown in \autoref{eq:construction:nu:fabrication}, any collection of $\vchannel$ functions can be composed such that they are equivalent to a $\vchannel$ that directly converts the input to the output. As with artists, $\vchannel$ are maps of sections such that the composition operators defined in \autoref{sec:artist:composition} can also act on transformers $\vchannel$, meaning that encoders can be added $\vchannel_{a+b} = \vchannel_{a} + \vchannel_{b}$ and multiplied d $\vchannel_{a\times b} = \vchannel_{a}  \vchannel_{b}$.  Encoders designed to satisfy these composability constraints provide for a rich set of building blocks for implementing complex encoders.

@@

    def draw(self, renderer, bounds, rate):
        for tau_local in self.section.query(bounds, rate): 
            mu = self.artist.nu(tau_local)
            rho = self.artist.qhat(**mu)
            output = rho(renderer)
\end{minted}
As shown in the \texttt{draw} method, generating a graphic section $\gsection$ is implemented as the composition of $\texttt{qhat} \approx \vmark$ and $\texttt{nu} \approx \vchannel$ applied to a local section of the sheaf $\texttt{self.section.query} \approx \desection^{i}$ such  $\texttt{draw} \approx \vmarkc\circ\vchannelc\circ\dsectionc  = \vartistc\circ \dsectionc$. The $\vchannel$ and $\vmark$ functions shown here are written such that they can generate a visual element given a local section $\dsection\restriction_{\dbase^{i}}$ which can be as little or large as needed. This flexibility is a prerequisite for building scalable and streaming visualizations that may not have access to all the data.

@@

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
\bibliography{IEEEabrv, bibliography}
